Sketch Recogniton using Particle Swarm Algorithm

### 2009-04-12 draft/chapterSurvey.tex
12-4

\chapter{Literature Survey}
\label{sec:survey}
% this will contians only the finished work .... and my comments to view the old survey see the file
%draft_chaptersurvey 
%\section{Introduction }
%\label{sec:Introduction}
The history of pen and sketch based computing goes back to 1963 with Ivan Sutherland's SKETCHPAD \cite{sutherlandsketch}, which used a light pen to draw on the monitor to create circuit diagrams. Rubine \cite{gestureexample12} Gesture recognition system. 
% i need to write a history then small introudction to the survey...
%  i have to state that the next  sections will generally group segmentation and symbol recognition task. 
\section{Preprocessing and Segmentation}
\label{sec:preprocessing}
%this section should describe how others did the first preprocessing of the problem. i have to summarize what i wrote before. 
 % first paragrah about how stroke is a messing thing and naturally it is ambigues. 
  In a system that supports free hand sketching, sketching should be as natural as possible. This means that the user can draw symbols using single or multiple strokes without constrains. The goal of a sketch system is making the user feel that using the using the system is nearly the same as using pencil and paper. But sketchs are incomplete drawings that are sloppy and messy, for example figures \ref{fig:Overshootandundershoot} and \ref{fig:overstroked} show examples of messy sketches drawn by users. This led researchers to perform sort of preprocessing before starting the sketch recognition task itself. The next sections explain different preprocessing techniques.

@@

One of the first systems in gestures recognition is Rubine's algorithm \cite{gestureexample12} . He used Gaussian statistical classifier to classify gestures.  Simple features as length of stroke, smoothness of stroke, the angle that the first part of the gesture is drawn at and properties of the bounding box of the stroke are used to train the classifier to recognize  symbols. In our system, we combine Rubine features with other to include the information on how the user draws sketch while recognizing the symbols. In \cite{Long00}  Long shows how similarities between gestures can be analyzed.  He used this information to identify gestures that are likely to confuse the recognition system. This helped interface designers to choose gestures that are easily recognized.

@@

 Systems that employ such representation must first preprocess strokes to break them into geometric primitives (see section \ref{sec:preprocessing} for a review on preprocessing methods). Recognition can then be treated as sub graph matching problem or constrain satisfaction problem.

@@

In the worst case, sub-graph matching is an exponential problem and thus is expensive to compute. Researcher tried to reduce complexity by restricting the search to fragments that are spatially or temporally grouped, or by using other assumption. Some of the common assumptions are that each stroke is only part of one symbol and that user must draw one shape before drawing the next\cite{physicalmeaning6}. These assumptions add constrains on how user can draw sketches and thus violates freely drawn sketches.  Moreover, the matching process in highly affected by segmentation results. Over or under fragmentation of stroke into geometric primitive complicates the matching process.  If a stroke is broken into too many fragments then the there will be extra component that will not map to any part of the symbol. If the stroke is divided enough then there will be component of the shape that are not filled while matching. The different ways of segmenting the sketches worsen the already high cost graph matching.

@@

@@

%Zernike moments were used by Hse et al. in \cite{}. 
Hse et al. \cite{HeloiseBeautification,zernike61} proposed using the magnitudes of Zernike moments as a shape property. Zernike moemnts are orthognals moments which define the distribution of point in the input space. The higher the order of the moment the finer the detail level the point distribution represents. Hse et al. \cite{HeloiseBeautification,zernike61} used a SVM classifier to train the features vector calculated from the moments. Early evluation of Zernike moemnts showed that they are invariant to both reflection and rotation of the input shape \cite{ }. In our system we used Zernike moemnts as part of our feature vector. A detailed comparison between Zernike moments and other features are demonstrated in Section \ref{}.

 Systems based on Global shape properties distinguish shapes by learning the different properties each shape class tend to have. These properties dose not depend on the order or number of strokes and thus are only based on apperance of the shape not how the user draw it. However, individual details of the shape are impossible to distinguish using these properties. For example, it is impoissible to use  to  differentiate between a current source symbole (circle contiaing and arrow Figure \ref{}) and a clock (circle containing a square signal Figure \ref{}).  Finear details of the symbols are not represented in this approach, thus shapes that have smalls details or only significant conceptual variation are not distiguished using this method. These symbols requires a system to represent different levels of shape details.

@@

In the third approach, the recognition foucus on appearance of shapes contrary to individual strokes or global shape properties. Kara \cite{imagetrainable48} used this approach while presenting a system that matches the input shape into a one of the  prototypes in the database. Figure \ref{fig:template} shows an example of prototypes of some of the shapes in the system. Input shapes are first pose and scale normalized then four image similarity measure are used to match input shapes to the database prototypes. He used two measures (Hausdor distance )  to measure similarity and two coeficients (Yule, Tanimoto) to measure dissimilarity. His approach allows the system to recongized a shape using only a single training sample as it is based on the shape prototype. However, the shape protoytpe must represent all  transformation and variation the input shape can have. Moreover, they restrict the image representation to 48x48 pixel due to computationa reasons. This restriction leads to elimintating some of the fine details that may be needed to distinguish some shape classes. There results on isolated symbols seems promising but the system have not beet tested on full sketch that tends to have high overlapping between symbols.

@@

In all of these system only \cite{4} using stroke based recognitiion have been implemented to the fully creat a system that free draw sketch and understand them. This is because the emphasis on the interactive side of the sketch sytem which result in large difficultes in processing free hand sketches. Methods based on gesture and global stroke properties produce systems that are dependent on assumptions about how strokes can be groubed. These systems assume that users will finish drawing a shape before moving to the next one \cite{6,11,14,15}. This let them to temporally group strokes and then recognize them based on their strokes or apperance properties. \cite{5} argued that this assumption dose not hold in natural sketchs.

Recogntion based on strokes using hierarchical shape description is applied in \cite{}. The principle problem in these systems is the computatnal complexity of the matching between shape description and geometeric primitves produced in the segmentaiton process. Natural sketch are dificult to segment and group to create reliable fragment of geometeric primitives. This is partially due to the increase of noise and other phenomena such as overlapped, overtraced strokes and multiple shapes drawn using

Natural sketches are dificult to reliably fragment into geometric primitives, because of the increase in noise and other phenomena such as interspersed drawing of parts from multiple shapes, multiple
shapes drawn with single strokes, and overtraced strokes. Because these systems are
dependent on the fragmentation process, they are either faced with performing recognition
with unreliable fragmentation results or must consider many possible ways to
fragment the sketch. Avoiding the fragmentation issue was one of our key motivating
factors in taking an appearance-based approach to recognition.
The appearance based methods described above have not been applied, to our
knowledge, to complete, freely drawn sketches. They have only been applied to
isolated shapes or synthetic images generated from isolated shapes. Although their
focus on the visual nature of the sketches appears promising for handling freely drawn
sketches they have not yet been evaluated on them.

@@

%review on the pso alogrithm alone.................

>>>>

\chapter{Literature Survey}
\label{sec:survey}
% this will contians only the finished work .... and my comments to view the old survey see the file
%draft_chaptersurvey 
%\section{Introduction }
%\label{sec:Introduction}
The history of pen and sketch based computing goes back to 1963 with Ivan Sutherland's SKETCHPAD \cite{sutherlandsketch}, which used a light pen to draw on the monitor to create circuit diagrams. Current technology and the increae in computuing power have increased the abilities of such interfaces and allowed the implementaion of fully sketching systems. Over the years, many other sketching based system have been developed, elthough throught this history most system foucused on building the interactive interfaces to assist users while entering information to the computer. To demonstrate our system inspiration we will first present current research in the area of sketch recognition. Fristly, preprosessing and segmentaiton methods will be discussed in details. Later, symbol recognition methods will be presented and analysied in Section \ref{sec:symbolrecogntion}.

% i need to write a history then small introudction to the survey...
%  i have to state that the next  sections will generally group segmentation and symbol recognition task. 
\section{Preprocessing and Segmentation}
\label{sec:preprocessing}
%this section should describe how others did the first preprocessing of the problem. i have to summarize what i wrote before. 
 % first paragrah about how stroke is a messing thing and naturally it is ambigues. 
  In a system that supports free hand sketching, sketching should be as natural as possible. This means that the user can draw symbols using single or multiple strokes without constrains. The goal of a sketch system is making the user feel that using the using the system is nearly the same as using pencil and paper. But sketchs are incomplete drawings that are sloppy and messy, for example figures \ref{fig:Overshootandundershoot} and \ref{fig:overstroked} show examples of messy sketches drawn by users. This led researchers to perform sort of preprocessing before starting the sketch recognition task itself. The next sections explain different preprocessing techniques.

@@

One of the first systems in gestures recognition is Rubine's algorithm \cite{gestureexample12} . He used Gaussian statistical classifier to classify gestures.  Simple features as length of stroke, smoothness of stroke, the angle that the first part of the gesture is drawn at and properties of the bounding box of the stroke are used to train the classifier to recognize  symbols. In our system, we combine Rubine features with others to include the information on how the user draws sketch while recognizing the symbols. In \cite{Long00}  Long shows how similarities between gestures can be analyzed.  He used this information to identify gestures that are likely to confuse the recognition system. This helped interface designers to choose gestures that are easily recognized.

@@

 Systems that employ such representation must first preprocess strokes to break them into geometric primitives (see section \ref{sec:preprocessing} for a review on preprocessing methods). Recognition can then be treated as sub graph matching problem.

@@

In the worst case, sub-graph matching is an exponential problem and thus is expensive to compute. Researcher tried to reduce complexity by restricting the search to fragments that are spatially or temporally grouped, or by using other assumption. Some of the common assumptions are that each stroke is only part of one symbol and that user must draw one shape before drawing the next\cite{physicalmeaning6}. These assumptions add constrains on how user can draw sketches and thus violates freely drawn sketches.  Moreover, the matching process in highly affected by segmentation results. Over or under fragmentation of stroke into geometric primitive complicates the matching process.  If a stroke is broken into too many fragments then there will be extra components that will not map to any part of the symbol. If the stroke is divided enough then there will be component of the shape that are not filled while matching. The different ways of segmenting the sketches worsen the already high cost graph matching.

@@

 \paragraph{Visual Describtive Languages:}
 \label{sec:VisualDescribtiveLanguages}

 Avoiding graph matching problems, research tried to use languages that discribe shapes instead of graphs. Visual describtive languages represents the shape into tokens of geometric primitives or relations between smaller  geometric primitives. Figures \ref{fig:arrow1} and \ref{fig:arrow2} show a regular arrow with 2 shafts and head and its definition in Ladder language \cite{Ladder30}. Shape recognition is then a question of parsing the sketch document into a recognizer that parse strokes according to the constructed grammar.

 Several visual grammer were introduced \cite{statisticalparsing26,Ladder30,GenericHMM28}. \cite{Ladder30,GenericHMM28}  Used visual description language which will generate the information of the structure and constrain of each object classified. The system  automatically generates java code for the recognizer for each object. The generated code cycles through the strokes drawn to check if any of the recognizers generated had interpreted any symbols partially or completely.  These systems are less computaitonal expensive than graph matching approach but they also operate based on the output of the segmentation process. Any error in the segmentation process lead to huge error while parsing the whole sketch. Moreover, symbols cannot share strokes between them which infleunce and restrict how the user draw symbols. 
 % complexity of the graph matching problem 
\begin{figure}
	\centering
		\subfigure[An Arrow]
			{\label{fig:arrow1}\includegraphics[scale=0.75]{../../neededfiles/Figures/arrowladderdef1.eps}}
			\hfill
		\subfigure[Arrow grammar definition]
		 {	\label{fig:arrow2}\includegraphics[scale=0.88]{../../neededfiles/Figures/arrowladderdef2.eps}}
	\caption[Arrow definition]
	\label{fig:arrowladderdef1}
\end{figure}

@@

%Zernike moments were used by Hse et al. in \cite{}. 
Hse et al. \cite{HeloiseBeautification,zernike61} proposed using the magnitudes of Zernike moments as a shape property. Zernike moemnts are orthognals moments which define the distribution of point in the input space. The higher the order of the moment the finer the detail level the point distribution represents. Hse et al. \cite{HeloiseBeautification,zernike61} used a SVM classifier to train the features vector calculated from the moments. Early evluation of Zernike moemnts showed that they are invariant to both reflection and rotation of the input shape \cite{ZerMomentOrthogonal}. In our system we used Zernike moemnts as part of our feature vector. A detailed comparison between Zernike moments and other features are demonstrated in Section \ref{sec:featuresComparisions}.

 \begin{figure}[]
	\centering
		\subfigure{Ac Source}{

		\includegraphics{images/acSource.eps}		\label{fig:acSource}}

 		\subfigure{Current Source}{

		\includegraphics{images/CurrentSource.eps} \label{fig:CurrentSource}}
	\caption{Example of current and ac sources}
	\label{fig:ACandCurrentSources}
\end{figure}

 Systems based on Global shape properties distinguish shapes by learning the different properties each shape class tend to have. These properties dose not depend on the order or number of strokes and thus are only based on apperance of the shape not how the user draw it. However, individual details of the shape are impossible to distinguish using these properties. For example, it is impoissible to use  to  differentiate between a current source symbole (circle contiaing and arrow Figure \ref{fig:CurrentSource}) and a Ac Source (circle containing a tilate Figure \ref{fig:acSource}).  Finear details of the symbols are not represented in this approach, thus shapes that have smalls details or only significant conceptual variation are not distiguished using this method. These symbols requires a system to represent different levels of shape details.

@@

In the third approach, the recognition foucus on appearance of shapes contrary to individual strokes or global shape properties. Kara \cite{imagetrainable48} used this approach while presenting a system that matches the input shape into a one of the  prototypes in the database. Figure \ref{fig:template} shows an example of prototypes of some of the shapes in the system. Input shapes are first pose and scale normalized then four image similarity measure are used to match input shapes to the database prototypes. He used two measures (Hausdor distance)  to measure similarity and two coeficients (Yule, Tanimoto) to measure dissimilarity. His approach allows the system to recongized a shape using only a single training sample as it is based on the shape prototype. However, the shape protoytpe must represent all  transformation and variation the input shape can have. Moreover, they restrict the image representation to 48x48 pixel due to computationa reasons. This restriction leads to elimintating some of the fine details that may be needed to distinguish some shape classes. There results on isolated symbols seems promising but the system have not beet tested on full sketch that tends to have high overlapping between symbols.

@@

In all of these system only \cite{SketchRead2007} using stroke based recognitiion have been implemented to the fully creat a system that free draw sketch and understand them. This is because the emphasis on the interactive side of the sketch sytem which result in large difficultes in processing free hand sketches. Methods based on gesture and global stroke properties produce systems that are dependent on assumptions about how strokes can be groubed. These systems assume that users will finish drawing a shape before moving to the next one \cite{6multistrokeEvaluation,Cali63,geometrydomain49}. This let them to temporally group strokes and then recognize them based on their strokes or apperance properties. \cite{AlvaradoDigital} argued that this assumption dose not hold in natural sketchs.

Recogntion based on strokes using hierarchical shape description is applied in \cite{SketchRead2007}. The principle problem in these systems is the computatnal complexity of the matching between shape description and geometeric primitves produced in the segmentaiton process. Natural sketch are dificult to segment and group to create reliable fragment of geometeric primitives. This is partially due to the increase of noise and other phenomena such as overlapped, overtraced strokes and sharing of strokes between different symbols.  Since these systems are depenable on the segmenation results, they either performe recognition using undependable segmentation results or consider different ways to segment the input sketch.

To avoid most problems in such system, our system uses hybird of both global shape and hierarchical discribtion methods. Currenlty, our system recognize on sysmbols at a time. Each symbol is segmented and hierarchical features are computed.  Features  that are represent the global discription of the shape are also generated and appended to the hierarchical featusres. The recognition is done using SVM Linear classifier as it proved to be one of the best learning algorithms.

%Natural sketches are dificult to reliably fragment into geometric primitives, because of the increase in noise and other phenomena such as interspersed drawing of parts from multiple shapes, multiple
%shapes drawn with single strokes, and overtraced strokes

 %The appearance based methods described above have not been applied, to our knowledge, to complete, freely drawn sketches. They have only been applied to isolated shapes or synthetic images generated from isolated shapes. Although their focus on the visual nature of the sketches appears promising for handling freely drawn sketches they have not yet been evaluated on them.

@@

%review on the pso alogrithm alone.................
%A literature overview of the simulation done on swarm behavior and there application in swarm robots. The survey  describe system that made biological modeling of birds, ants behavior. Birds was modeled in "BIOD" and Flack systems which only modeled the distributed behaviors of flock of birds.  Kube and zhang simulated the behavior of ants in his system. Tersopooulos et al. developed artificial fish in a 3d physical world that emulate real fish behavior. \\

The survey then mention swarm robots and the literature over it. Beoi proposed a swarm cellar of autonomously  robots. others  introduced a behavior based swarms  Beni also proposed a sign post system where robots are sample but act under the influence of "sign post robot" which change state of the robot by passing by a "sign post robot".  other studied the affect of communication between robot and it was found that  as inter robot comunication improve more complex behavior can be achieved. \subsection*{A Hybrid network for feature extraction}
The paper uses the SIDS to generate correct$x_{i}$ and $x_{j}$t mapping of features in images. its is divided into two steps. First FCS feature classifier system based on simple one layer neural networks. Each cell compute a micro function of object to be classified. The system use an n tuple feature classification  . The second is a feature search system that  use to search the n tuple mapping if feature is present in this mapping.

%\section{Paper:89  [2005]
}
\subsection*{Dynamic probabilistic particle swarms}
PSO is a swarm optimization where the particle are moved around the search space using direction and velocity equation. Each particle have location equation x and V velocity equation to compute next location in the solution space. The basic algorithm use constants or static tuning parameters with the help of particle best and global best to influence the equations of particle transformation. The paper present a dynamic probabilistic based on more probability and  information than the original algorithm.\\

There are two proposals first is trimmed uniform probabilities where new positions of particle is computed on biases of expected position of old best location of the particle and global best. The second is Gaussian  probabilities where the influence of moving particles are give based on the probability of how near the particle to the solution X. the propability is computed using gaussian distribution and subsituted to influence in the transformation equation of the particels.

%\section{Paper:90  [2000]}
\subsection*{Artificial Ant colonies in digital image habitats a mass behavior effect study on pattern recognition}
The paper describe how ant can be used in image processing environments. The system will need to have describe how ant move around the image using the transition rules from one location to another.  The transition rule $w(\sigma)=\left( 1+\frac{\sigma}{1+\sigma\delta}\right)^{\beta} $ Function depend on the pheromone density on the environment of the pixel and neighborhood pixels. The system also include a definition of the rate $\eta$ that the ant emit pheromonal and the rate \textit{k }the rate pheromal evaporate. 
%\section{Paper:91  [2004]}
\subsection*{An algorithm for swarm based color image segmentation}
The Ant system was used to segment color images using a system that is described in this paper. The algorithm main goal was to segment pixel of similar color into segments. The paper describe in details the data Structure and the swarm structure of the system.\\

The Data is located in 2D matrices which represent the environment [pixel info] and the segmented map [what ants write data in] each with the same size as the original image. There is also a Color map array that is used as a label for the segment map.\\

The swarm system is divided into to main agents Scots and worker. The Scots are the agent that have a sense of color based on an equation and error bound. The Scot loop over color label form the color map move around the image randomly till he found its found a similar color.  When it found a similar color it write to the segment map the  label in the pixel location then check the neighbors to see if they have a similar color.  If not then the Scot is moved around into different color and location. The worker work on the segment map where intailly  it has a fixed color liable and its moves randomly on location till it found a segmented location with same label as itself. The worker check the neighborhood location if the Scots has label it or not. if it is labeled with different color then the larger of the two labeled will still be present else  moves randomly till found an interesting neighborhood.
